Do investors in clean energy ETFs herd?

### 2025-01-02 alt_energy_herding.qmd
positron test
improved introduction and literature review

---
format:
  pdf:
    tbl-cap-location: top
    dpi: 500
    fig-pos: "H"
    number-sections: true
    cite-method: natbib
    include-in-header: 
      - text: |
          \usepackage{lscape}
          \newcommand{\blandscape}{\begin{landscape}}
          \newcommand{\elandscape}{\end{landscape}}
          \usepackage{rotating}
          \newcommand{\bsideways}{\begin{sidewaystable}[htbp]}
          \newcommand{\esideways}{\end{sidewaystable}}
      - file: preamble.tex 
    include-before-body: title.tex
    keep-tex: TRUE
  docx:
    reference-doc: new_template_no_numbering.docx
pdf-engine: lualatex
# monofont: "DejaVu Sans Mono"
# mainfont: "Times New Roman"
execute: 
  echo: false
  warning: false
editor: source
bibliography: references.bib
biblio-style: apalike
natbiboptions: round
link-citations: true
linkcolor: blue!100!black
urlcolor: blue!100!black
citecolor: blue!100!black
linestretch: 1.5
indent: TRUE
title-block-banner: false
title-block-style: none
---

@@

# Introduction

Climate sustainability has become a major concern for financial markets and investors. This growing interest is primary driven by the increasing materialisation of climate risks, and the actions of governments, institutions and organizations towards a sustainable future [@giglio2021a]. In order to ensure returns investors increasingly seek to herd against climate risks by investing in green financial products. Although the evidence is mixed, indications that returns from green financial products are comparable to traditional financial [see amongst other, @decclesia2024; @nguyen2025; @pastor2022; and @naqvi2022]. Therefore, the growth of green assets-under-management is likely to continue.

A key feature of the green financial products is the exchange traded funds (ETFs). ETFs are a type of security that involves a collection of securities that often tracks an underlying index, although they can invest in any number of industry sectors or use various strategies. In addition, environment, social, and governance (ESG) ETFs serve as a market discovery tool for investors to identify and invest in companies that are environmentally friendly [@briere2023]. Thus, ESG ETFs exhibit environment friendly characteristics.

Among the ESG ETFs, the Clean Energy (CE) ETFs have been the best-performing ones in 2022, followed by the Cybersecurity and Artificial Intelligence (AI) ETFs [@decclesia2024]. The clean energy transition represents one of the largest multi-decade secular growth opportunities.  After the inclusion of Green energy financing in the list of United Nations Sustainability Goals (SDGs) as SDG 7, the role, importance, and visibility of green financial products; all have escalated enormously [@naqvi2022].

@@

Following the relevant literature [@christie1995; and @chang2000], we compute dispersion of the $ith$ ETF from the market return. This is known as the Cross Sectional Absolute Deviation  ($CSAD_t$) measure. Empirically the $CSAD_t$ is defined in the following manner:

>>>>

---
format:
  pdf:
    tbl-cap-location: top
    dpi: 500
    fig-pos: "H"
    number-sections: true
    cite-method: natbib
    include-in-header: 
      - text: |
          \usepackage{lscape}
          \newcommand{\blandscape}{\begin{landscape}}
          \newcommand{\elandscape}{\end{landscape}}
          \usepackage{rotating}
          \newcommand{\bsideways}{\begin{sidewaystable}[htbp]}
          \newcommand{\esideways}{\end{sidewaystable}}
      - file: preamble.tex 
    include-before-body: title.tex
    keep-tex: TRUE
  docx:
    reference-doc: new_template_no_numbering.docx
pdf-engine: lualatex
# monofont: "DejaVu Sans Mono"
# mainfont: "Times New Roman"
execute: 
  echo: false
  warning: false
editor: source
bibliography: references.bib
biblio-style: apalike
natbiboptions: round
link-citations: true
linkcolor: blue!100!black
urlcolor: blue!100!black
citecolor: blue!100!black
linestretch: 1.5
indent: TRUE
title-block-banner: false
title-block-style: none
render-on-save: true
---

@@

# Introduction and literature review

Climate sustainability has become a major concern for financial markets and investors. This growing interest is primary driven by the increasing materialisation of climate risks, and the actions of governments, institutions and organizations towards a sustainable future [@giglio2021a]. In order to ensure returns, investors increasingly seek to hedge against climate risks by investing in green financial products. Although there evidence is mixed, indications that returns from green financial products are comparable to traditional financial [see amongst others, @decclesia2024; @nguyen2025; @pastor2022; and @naqvi2022].

Specifically, climate change present risk to investor portfolios through two primary sources, physical and transition risks. physical risks or direct impact refer to extreme climate events such as floods and droughts, which impact business operations and infrastructure; and transition risks are the policy, technological, and other costs that societies bear to achieve low carbon economies [@nguyen2025; and @giglio2021a amongst others]. Investors, therefore, recognise these risk and seek to mitigate against them, as they seek return enhancing green financial products.

Exchange traded funds (ETFs) are a key feature of green financial products. ETFs are a type of security that involves a collection of securities that often tracks an underlying index, although they can invest in any number of industry sectors or use various strategies. In addition, environment, social, and governance (ESG) ETFs serve as a market discovery tool for investors to identify and invest in companies that are environmentally friendly [@briere2023]. Among the ESG ETFs, the Clean Energy (CE) ETFs have been the best-performing ones in 2022, followed by the Cybersecurity and Artificial Intelligence (AI) ETFs [@decclesia2024]. The clean energy transition represents one of the largest multi-decade secular growth opportunities.  After the inclusion of Green energy financing in the list of United Nations Sustainability Goals (SDGs) as SDG 7, the role, importance, and visibility of green financial products; all have escalated enormously [@naqvi2022]. Therefore, the growth of green assets-under-management is likely to continue.

\newpage
In this new climate sustainability paradigm, investors face many pressures that not only bear on returns, but also on the stability of financial markets. For example, resulting regulations aimed at reducing emissions can surprisingly reduce the profitability of fossil-fuel based companies, or the possible mispricing of assets from ignoring climate risks can lead to significant losses [@nguyen2025]. In addition to these climate risks, a general change in investor attitudes can drive the inclusion of green assets in their portfolio leading to systemic risk. These pressures are not withstanding the possible contribution that financial markets can play in mitigating and reducing the negative effects of climate change [@giglio2021a].

However, given how recent the inclusion of climate sustainability investment decisions is, it is not clear what the actual impact will be in the long run. In this study specifically, we ask whether the rapid adoption of CE ETFs could be driven a market fad driven by herding behaviour, or is a fundamental change in investor behaviour? Investors, for example, can believe that peers have information about climate risks that they do not, investors may herd to avoid losses compared to peers, or investors may be encouraged to herd by the desire to align to climate related social values [@ciciretti2021; @gavrilakis2023; @loang2023]. Therefore, market volatility and crisis, financial performance, and investor sentiment can be drivers of herding in ESG markets

Although the overall evidence is mixed, herding behaviour in ESG markets is not without precedent. Amongst others, @loang2023 found that compliance with SDG goals can introduce bias in investor sentiment, which lead to herding behaviour. Using a Twitter (or X) uncertainty index, @koutmos2024 found evidence of herding in US-based ESG index fund investors. @przychodzen2016 found herding behaviour amongst fund managers who incorporated ESG strategies in their portfolios. Lastly, @rubbaniy2021 highlight evidence of herding in the MSCI US ESG Leader Index during extreme periods (bear and bull periods).

@@

Following the relevant literature [@christie1995; and @chang2000], we compute dispersion of the $i^{th}$ ETF from the market return. This is known as the Cross Sectional Absolute Deviation  ($CSAD_t$) measure. Empirically the $CSAD_t$ is defined in the following manner:

### 2025-01-02 alt_energy_herding.tex
improved introduction and literature review

\setstretch{1.5}
\section{Introduction}\label{introduction}

Climate sustainability has become a major concern for financial markets
and investors. This growing interest is primary driven by the increasing
materialisation of climate risks, and the actions of governments,
institutions and organizations towards a sustainable future
\citep{giglio2021a}. In order to ensure returns investors increasingly
seek to herd against climate risks by investing in green financial
products. Although the evidence is mixed, indications that returns from
green financial products are comparable to traditional financial
\citetext{\citealp[see amongst
other,][]{decclesia2024}; \citealp{nguyen2025}; \citealp{pastor2022}; \citealp[and][]{naqvi2022}}.
Therefore, the growth of green assets-under-management is likely to
continue.

A key feature of the green financial products is the exchange traded
funds (ETFs). ETFs are a type of security that involves a collection of
securities that often tracks an underlying index, although they can
invest in any number of industry sectors or use various strategies. In
addition, environment, social, and governance (ESG) ETFs serve as a
market discovery tool for investors to identify and invest in companies
that are environmentally friendly \citep{briere2023}. Thus, ESG ETFs
exhibit environment friendly characteristics.

Among the ESG ETFs, the Clean Energy (CE) ETFs have been the
best-performing ones in 2022, followed by the Cybersecurity and
Artificial Intelligence (AI) ETFs \citep{decclesia2024}. The clean
energy transition represents one of the largest multi-decade secular
growth opportunities. After the inclusion of Green energy financing in
the list of United Nations Sustainability Goals (SDGs) as SDG 7, the
role, importance, and visibility of green financial products; all have
escalated enormously \citep{naqvi2022}.

@@

Following the relevant literature
\citetext{\citealp{christie1995}; \citealp[and][]{chang2000}}, we
compute dispersion of the \(ith\) ETF from the market return. This is
known as the Cross Sectional Absolute Deviation (\(CSAD_t\)) measure.
Empirically the \(CSAD_t\) is defined in the following manner:

>>>>

\setstretch{1.5}
\section{Introduction and literature
review}\label{introduction-and-literature-review}

Climate sustainability has become a major concern for financial markets
and investors. This growing interest is primary driven by the increasing
materialisation of climate risks, and the actions of governments,
institutions and organizations towards a sustainable future
\citep{giglio2021a}. In order to ensure returns, investors increasingly
seek to hedge against climate risks by investing in green financial
products. Although there evidence is mixed, indications that returns
from green financial products are comparable to traditional financial
\citetext{\citealp[see amongst
others,][]{decclesia2024}; \citealp{nguyen2025}; \citealp{pastor2022}; \citealp[and][]{naqvi2022}}.

Specifically, climate change present risk to investor portfolios through
two primary sources, physical and transition risks. physical risks or
direct impact refer to extreme climate events such as floods and
droughts, which impact business operations and infrastructure; and
transition risks are the policy, technological, and other costs that
societies bear to achieve low carbon economies
\citetext{\citealp{nguyen2025}; \citealp[and][ amongst
others]{giglio2021a}}. Investors, therefore, recognise these risk and
seek to mitigate against them, as they seek return enhancing green
financial products.

Exchange traded funds (ETFs) are a key feature of green financial
products. ETFs are a type of security that involves a collection of
securities that often tracks an underlying index, although they can
invest in any number of industry sectors or use various strategies. In
addition, environment, social, and governance (ESG) ETFs serve as a
market discovery tool for investors to identify and invest in companies
that are environmentally friendly \citep{briere2023}. Among the ESG
ETFs, the Clean Energy (CE) ETFs have been the best-performing ones in
2022, followed by the Cybersecurity and Artificial Intelligence (AI)
ETFs \citep{decclesia2024}. The clean energy transition represents one
of the largest multi-decade secular growth opportunities. After the
inclusion of Green energy financing in the list of United Nations
Sustainability Goals (SDGs) as SDG 7, the role, importance, and
visibility of green financial products; all have escalated enormously
\citep{naqvi2022}. Therefore, the growth of green
assets-under-management is likely to continue.

In this new climate sustainability paradigm, investors face many
pressures that not only bear on returns, but also on the stability of
financial markets. For example, resulting regulations aimed at reducing
emissions can surprisingly reduce the profitability of fossil-fuel based
companies, or the possible mispricing of assets from ignoring climate
risks can lead to significant losses \citep{nguyen2025}. In addition to
these climate risks, a general change in investor attitudes can drive
the inclusion of green assets in their portfolio leading to systemic
risk. These pressures are not withstanding the possible contribution
that financial markets can play in mitigating and reducing the negative
effects of climate change \citep{giglio2021a}.

However, given how recent the inclusion of climate sustainability
investment decisions is, it is not clear what the actual impact will be
in the long run. In this study specifically, we ask whether the rapid
adoption of CE ETFs could be driven a market fad driven by herding
behaviour, or is a fundamental change in investor behaviour? Investors,
for example, can believe that peers have information about climate risks
that they do not, investors may herd to avoid losses compared to peers,
or investors may be encouraged to herd by the desire to align to climate
related social values \citep{ciciretti2021, gavrilakis2023, loang2023}.
Therefore, market volatility and crisis, financial performance, and
investor sentiment can be drivers of herding in ESG markets

Although the overall evidence is mixed, herding behaviour in ESG markets
is not without precedent. Amongst others, \citet{loang2023} found that
compliance with SDG goals can introduce bias in investor sentiment,
which lead to herding behaviour. Using a Twitter (or X) uncertainty
index, \citet{koutmos2024} found evidence of herding in US-based ESG
index fund investors. \citet{przychodzen2016} found herding behaviour
amongst fund managers who incorporated ESG strategies in their
portfolios. Lastly, \citet{rubbaniy2021} highlight evidence of herding
in the MSCI US ESG Leader Index during extreme periods (bear and bull
periods).

@@

Following the relevant literature
\citetext{\citealp{christie1995}; \citealp[and][]{chang2000}}, we
compute dispersion of the \(i^{th}\) ETF from the market return. This is
known as the Cross Sectional Absolute Deviation (\(CSAD_t\)) measure.
Empirically the \(CSAD_t\) is defined in the following manner:
